---
title: "Examining the Moderating Effect of Rejection Sensitivity on the Relationship between Discrimination and Telomere Length in African American Women"
author: "Amanda D. Perez - UC Berkeley, Tyan Parker-Dominguez - USC, Amani M. Allen - UC Berkeley"
format:
  html:
    toc: true
    theme: yeti
---

```{r, include=FALSE, warning = F, message = F}
library(lavaan)
library(tidyverse)
library(psych)
library(kableExtra)
library(haven)
library(readxl)
library(gridExtra)
library(misty)
library(apaTables)
library(jtools)
library(interactions)
library(semTools)
library(mice)

load("C:/Users/Admin/Dropbox/HEARTs/RS/Imputations/2023-03-01_imputed-aawhhs-data-rs.RData")
```

## Abstract

::: callout-note
## **Rejection sensitivity** intensifies the negative effect of racial discrimination on **physiological aging**
:::

Rejection sensitivity is a learned trait, formed in part from past experiences of prejudice and discrimination, that refers to the tendency of individuals to anxiously expect, readily perceive, and intensely respond to cues of rejection in social interactions. Previous research has linked higher levels of rejection sensitivity to negative health outcomes such as increased symptoms of anxiety and depression, inflammation, and higher cortisol levels. Using a nationally representative sample of 207 African American women, we examined the moderating effect of race-based and gender-based rejection sensitivity on the relationship between experiences of racial discrimination and telomere length, a marker of cellular aging and overall health. Racial discrimination was measured using the 8-item experiences of discrimination scale (ð›¼=.91). Race- and gender-based rejection sensitivity were measured using the 6 item race-based rejection sensitivity scale (ð›¼=.91) and 8 item gender-based rejection sensitivity scale (ð›¼=.94), respectively. We found that rejection sensitivity, whether race-based (Î²=-0.25, p=0.045 ) or gender-based (Î²=-0.21, p=0.03), moderated the relationship between experiences of discrimination and telomere length. Individuals reporting low levels of either types of rejection sensitivity, and who also reported higher vs. lower racial discrimination had the longest telomere lengths. These findings highlight the importance of considering rejection sensitivity as a factor that can exacerbate the negative effects of discrimination on health outcomes, particularly in African American women who are at a higher risk for discrimination and related health disparities.

*Contact: adpc@berkeley.edu*

## Tables

**Demographic Table**

|                    |           |           |
|--------------------|-----------|-----------|
| Term               | N         | \%        |
| **Age (M, SD)**    | **41.72** | **5.90**  |
| **\> HS Diploma**  | **138**   | **66.67** |
| **Not Employed**   | **93**    | **44.93** |
| **Not Married**    | **146**   | **70.53** |
| **Taking CV Meds** | **12**    | **5.80**  |

**Exposures and Outcomes**

|                                        |           |          |             |
|----------------------------------------|-----------|----------|-------------|
| Term                                   | Mean      | SD       | Range       |
| **Gender-based Rejection Sensitivity** | **10.67** | **8.04** | **1-32.11** |
| **Race-based Rejection Sensitivity**   | **10.03** | **8.70** | **1-31**    |
| **Experiences of Discrimination**      | **19.56** | **9.18** | **8-40**    |
| **Neuroticism**                        | **3.08**  | **0.77** | **1-5**     |

```{r, warning = F, message = F, include = F}
#### Data Cleaning

# centering variables per Yije's code
dta_long$eod_sum_c <- misty::center(dta_long$eod_sum, type = "CWC", cluster = dta_long$.imp)
dta_long$rs_gender_c <- misty::center(dta_long$GBRS_avg,  type = "CWC", cluster = dta_long$.imp)
dta_long$rs_race_c <- misty::center(dta_long$RBRS_avg, type = "CWC", cluster  = dta_long$.imp)
dta_long$sc_race_c <- misty::center(dta_long$sc_race_avg, type = "CWC", cluster = dta_long$.imp)
dta_long$sc_gender_c <- misty::center(dta_long$sc_gender_avg, type = "CWC", cluster = dta_long$.imp)
dta_long$rrw_c <- misty::center(dta_long$art_rrw, type = "CWC", cluster = dta_long$.imp)
dta_long$sa_c <- misty::center(dta_long$art_stereotype_awareness, type = "CWC", cluster = dta_long$.imp)
dta_long$cc_c <- misty::center(dta_long$art_confirmation_concern, type = "CWC", cluster = dta_long$.imp)


dta_long$eod2 <- dta_long$eod_sum_c * dta_long$eod_sum_c

dta_long$cvmeds_current <- misty::center(dta_long$cvmeds_current, type = "CWC", cluster = dta_long$.imp)
dta_long$diabetesmedtx_current <- misty::center(dta_long$diabetesmedtx_current, type = "CWC", cluster = dta_long$.imp)
dta_long$educ_dichot_c <- misty::center(dta_long$educ_dichot, type = "CWC", cluster = dta_long$.imp)
dta_long$pov_c <- misty::center(dta_long$pov, type = "CWC", cluster = dta_long$.imp)
dta_long$age_yrs_c <- misty::center(dta_long$age_yrs, type = "CWC", cluster = dta_long$.imp)
dta_long$neurot_avg_c <- misty::center(dta_long$neurot_avg, type = "CWC", cluster = dta_long$.imp)
dta_long$marstatus_dichot_c <- misty::center(dta_long$marstatus_dichot, type = "CWC", cluster = dta_long$.imp)
dta_long$employ_dichot_c <- misty::center(dta_long$employ_dichot, type = "CWC", cluster = dta_long$.imp)

# creating interactions
dta_long$int_rs_sa <- dta_long$rs_race_c * dta_long$sa_c
dta_long$int_rs_cc <- dta_long$rs_race_c * dta_long$cc_c
dta_long$int_rs_rrw <- dta_long$rs_race_c * dta_long$rrw_c
dta_long$int_rs_sc_race <- dta_long$rs_race_c * dta_long$sc_race_c
dta_long$int_rs_sc_gender <- dta_long$sc_gender_c * dta_long$rs_gender_c
dta_long$int_rs <- dta_long$rs_race_c * dta_long$rs_gender_c
dta_long$int_race <- dta_long$rs_race_c * dta_long$eod_sum_c
dta_long$int_gender <- dta_long$eod_sum_c * dta_long$rs_gender_c
dta_long$int_race2 <- dta_long$rs_race_c * dta_long$eod2
dta_long$int_gender2 <- dta_long$eod2 * dta_long$rs_gender_c

#dta_long <- dta_long[dta_long$.id != 184, ] #removing outlier 
```

```{r, warning = F, message = F, include = F}
#dta_long <- as.data.frame(dta_long)
dta_mids <- as.mids(dta_long, .imp = ".imp", .id = ".id")
mice.imp <- NULL
for(i in 1:20) {
  mice.imp[[i]] <- mice::complete(dta_mids, action = i, inc=FALSE)  }
#saveRDS(mice.imp, "mice.complete.rs.full.rds")
```

## Visuals

```{r, echo = FALSE}
#windowsFonts()
windowsFonts("TT Times New Roman" = windowsFont("TT Times New Roman"))

##custom theme
theme_perez <- function() {
    ggplot2::theme_bw() +
    ggplot2::theme(
      axis.text = element_text(colour = "black",  family = "TT Times New Roman", size = 21),
    axis.title = element_text(colour = "black", family = "TT Times New Roman", size = 23),
    legend.title = element_text(colour = "black", family = "TT Times New Roman", size = 23),
    legend.text = element_text(colour = "black", family = "TT Times New Roman", size = 23),
    plot.title = element_text(colour = "black", family = "TT Times New Roman", size = 23,face = "bold"),
    plot.subtitle = element_text(colour = "black", family = "TT Times New Roman", size = 23),
    legend.position="right"
      )
}

```

```{r, echo = FALSE}
library(RColorBrewer)
pal <- c("orange", "#0c2c84")
```

```{r, echo = FALSE}
fiti  <- lm(tlr ~ eod_avg * RBRS_avg + cvmeds_current + diabetesmedtx_current + educ_dichot_c + pov_c + age_yrs_c + marstatus_dichot_c + employ_dichot_c + neurot_avg_c + sc_race_c, mice.imp[[1]])

interact_plot(fiti, modx = RBRS_avg, pred = eod_avg, modx.values = "plus-minus", x.label = "EOD", y.label = "rTL", modx.labels = c("Low", "High"), legend.main = "Race-Based RS", colors = pal) + theme_perez()
```

```{r, echo = FALSE}
fiti  <- lm(tlr ~ eod_avg * GBRS_avg + cvmeds_current + diabetesmedtx_current + educ_dichot_c + pov_c + age_yrs_c + marstatus_dichot_c + employ_dichot_c + neurot_avg_c + sc_gender_c, mice.imp[[1]])

interact_plot(fiti, modx = GBRS_avg, pred = eod_avg, modx.values = "plus-minus", x.label = "EOD", y.label = "rTL", modx.labels = c("Low", "High"), legend.main = "Gender-Based RS",colors = pal) + theme_perez()
```
